---
title: "ESM 206 A3 Task2"
author: "Hollie Pennington"
date: "11/4/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE, warning = FALSE, echo = FALSE, messages = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center')
#install packages
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
```

# Introduction: 
The report shows a comparison between of weight for juvenile snowshoe hares, Lepus americanus. Weight is compared by sex and site and is compared to the size of hind feet. Samples were collected in Alaska from the Tanana valley between the years 1999 and 2012. Weight is compared by site and sex (Kiellend et al. 2017). 

Data and analyses: The data in this exploratory report was accumulated by Kiellend et al. 2017. To analyse and compare data I used bar graphs, box plots, and linear regression visuals, along with exploratory histograms and qq plots that are not depicted. To find statistical relevance I tested for means, standard deviations, R2, Cohen's D, unpaired two-sample t-tests, and Pearson's r correlation. 

## Juvenile Hare Initial Investigation 

```{r}
snowshoe <- read_csv(here("data","bonanza_hares.csv")) %>% 
  clean_names() %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date))
#read in data 
#made a year column using mutate
```

```{r}
j_snowshoe <- snowshoe %>% 
  filter(age == "j") %>% 
  group_by(age, year) %>% 
  summarise(total = n()) 
#created new data frame including only juvenile hares and they are grouped by age and year. 
```

```{r}
ggplot(data = j_snowshoe, aes(x = year, y = total)) + 
  geom_col(fill = "blue", color = "black") + 
  labs(title = "Annual juvenile hare trap count", x = "Year", y = "Total number of juvenile hares") + 
  theme_test()
#created bar graph comparing year to the number of hares trapped 
```

<b> Figure 1: </b> This graph above represents the total number of juvenile hare trappings from year 1998 to 2012. The year with the most trappings is 1999. The year with the least trappings is 2010. The graph above suggest a declining population of juvenile snowshoe hares.      

Takeaways: The minimum amount of snowshoes that were found in one single year was 2. The highest number of snowshoes found in one year was 126. The median amount of snowshoes being trapped throughout the years is 18.5. The mean amount of hares trapped every year was 12. I suspect that there are confounding factors such as effort, number of traps, and hare location at play that impacted the data. In future studies the confounding factors can be reduced by making sure all human efforts and methods are consistent throughout the years. 


```{r}
j_snowshoe_stats <- j_snowshoe %>% 
  summarise(minimum = min(total), 
            maximum = max(total), 
            mean = mean(total, ra.nm = TRUE), 
            median = median(total, ra.nm = TRUE), 
            total = n())
#found stats for variables with juvenile hare data frame 
```

## Juvenile Hare Weights 

```{r}
weights_j_snowshoe <- snowshoe %>% 
  filter(age == "j") %>% 
  select(age, weight, grid, sex) %>% 
   mutate(sex = case_when(
    sex == "f" ~ "Female", 
    sex == "m" ~ "Male",
    TRUE ~ "Unknown")) %>% 
  mutate(grid = case_when(
    grid == "bonbs" ~ "Bonbs", 
    grid == "bonmat" ~ "Bonmat", 
    grid == "bonrip" ~ "Bonrip"
  )) %>% #changed names 
  group_by(sex, weight, grid, age) %>% 
  summarise(total = n())
#created new data frame for juvenile hare that includes age, weight, grid, and sex
```

```{r, include = FALSE}
#Attempted visualization, did not like graph, not including in knitted document but I worked really hard on this and don't want to delete it in case I need the reference later
ggplot(data = weights_j_snowshoe, aes(y = weight, x = total, fill = sex)) + 
  geom_col(position = position_dodge()) + #created graph
  facet_wrap(~ grid) + #faceted by grid
  labs(title = "Juvenile snowshoe weight by site", 
       x = "Number of hares", 
       y = "Hare weight (g)",
       fill = "Sex") +
  theme(legend.position = c(0.9, 0.85), 
        legend.key.size = unit(0.3, 'cm'), 
        legend.text = element_text(size = 7), 
        legend.title = element_text(size = 8)) #changed graph visuals 
```

```{r}
ggplot(data = weights_j_snowshoe, aes(x = sex, y = weight, fill = sex)) + 
  geom_boxplot(show.legend = FALSE) + #created graph
  facet_wrap(~ grid) + #faceted by grid
  labs(title = "Comparing juvenile snowshoe weight and by site",
       x = "Sex", 
       y = "Weight (g)")+
  stat_summary(show.legend = FALSE) +
  theme_test()
```

<b> FIGURE 2: </b> The above box plot represents hare weight (g) compared to hare sex. The data is divided by the three different sites Bonbs, Bonmat, Bonrip. The large black dots represent the mean value along with the standard deviation. 

Takeaways: The median weight for male snowshoe hares in larger than the median weight for female snowshoe hares. In the Bonbs site and the Bonmat site there is a larger variance in female weight then there is in male weight. Female hares tend to have a more evenly distribute weight. 
 
## Juvenile hare weight comparison (male & female snowshoe hares)


```{r}
stats_snowshoe <- weights_j_snowshoe %>% 
  select(sex) %>% 
  group_by(sex) %>% 
  summarise(mean = mean(weight, na.rm = TRUE),
            standard_deviation = sd(weight, na.rm = TRUE),
            sample_size = n())
#created new data frame with statistics on j hare weights 
```

```{r}
stats_snowshoe %>% 
kable(col.name = c("Sex", "Mean Weight (g)", "Standard Deviation (g)", "Sample Size"), #created graph
      caption = "<b> Table 1: </b> The table below shows the difference between mean female hare weight (g) and mean male hare weight(g). Males have a higher mean weight by 81.63 grams.") %>%  #added caption 
  kable_styling(full_width = FALSE)
#created visual on snowshoe statistics 
```

```{r, include = FALSE}
ggplot(data = weights_j_snowshoe, aes(x = weight)) + 
  geom_histogram(bins = 15) + 
  facet_wrap(~ sex)

ggplot(data = weights_j_snowshoe, aes(sample = weight)) + 
  geom_qq() + 
  facet_wrap(~ sex)
#visualize data to see if we can run statistical testing. 
#based on outcome of graphs and number of samples, I feel confident that I can do a t-test. 
```

```{r, include = FALSE}
f_hare <- weights_j_snowshoe %>% 
  filter(sex == "Female") %>% 
  pull(weight)

m_hare <- weights_j_snowshoe %>% 
  filter(sex == "Male") %>% 
  pull(weight)

ttest_j_hare <- t.test(f_hare, 
                       m_hare)
ttest_j_hare

cd_j_hare <- cohen.d(f_hare,
                     m_hare, 
                     na.rm = TRUE)

cd_j_hare
```
 
Takeaways: The difference in means between the female hares and males hares is `r round(ttest_j_hare$estimate[1],2)` g,  with the male hares having the larger mean. The effect size is very small (Cohen's d = `r round(cd_j_hare$estimate,2)`) and difference significant. There is a less than 5% chance that the samples were drawn from populations with different sizes (P-value = 0.049).

## Relationship between juvenile weight and hind foot length 

```{r}
j_snowshoes <- snowshoe %>% 
  filter(age == "j") %>% 
  group_by(weight, hindft) %>% 
  summarise(total = n())
#created new data frame to help compare weight by hind feet
```

```{r}
#relationship between juvenile weight and hind foot length
ggplot(data = j_snowshoes, aes(y = weight, x = hindft)) + 
  geom_jitter() + #created graph
  geom_smooth(method = lm) + #added tred line
  labs(title = "Juvenile hares weight vs hind feet", x = "Hind feet length (mm)", y = "Weight (g)") + #edited graph labels
  theme_test()
```

<b> Figure 3: </b> The above graph shows a positive correlation between weight (g) and hind feet (mm) in juvenile snowshoe hares. 

## Linear regression: 

```{r, include = FALSE}
#linear regression 
lm_j_snowshoes <- lm(hindft ~ weight, data = j_snowshoes)
lm_j_snowshoes
summary(lm_j_snowshoes)

glance_lm_j_snowshoes <- glance(lm_j_snowshoes)
tidy_lm_j_snowshoes <- tidy(lm_j_snowshoes)

cor_j_hare <- cor.test(j_snowshoes$weight, j_snowshoes$hindft)

cor_j_hare

tidy_cor_j_hare <- tidy(cor_j_hare)
```

Takeaways:  The relationship between weight(g) and hind feet length (mm) is explored across juvenile hares across all three sites. Further analysis can be completed on the relationship between sex, weight, and hind feet length. 

The relationship between weight (g) and hind feet length (mm) explored in figure three reveals positive linear relationship, indicating that weight (g) and hind feet length (mm) and correlated. Simple linear regressions shows that hind feet length significantly predicts snowshoe hare weight (g) (p < 0.001,  R^2^ = `r round(glance_lm_j_snowshoes$r.squared,3)`)
with an average slope of $\beta$ = `r round(tidy_lm_j_snowshoes$estimate[2], 3)` grams. For each one millimeter in hind feet we can expect on average an increase of weight in grams of `r round(tidy_lm_j_snowshoes$estimate[2], 3)` (g). Weight and hind feet length are significantly and positively correlated (Person's *r* =`r round(tidy_cor_j_hare$estimate,3)`,
p < 0.001). Diagnostic plots (not included) reveal relatively normally distributed and heteroscedastic residuals. 


```{r, include = FALSE}
plot(lm_j_snowshoes)
```

## Summary: 
The initial findings of this report reveal:  

- There is a decline in population of snow shoe hares.

- Male snowshoe hares on average have a larger weight than female snowshoe hares. 

- There is a positive correlation between weight and hind feet. 

### Citation: 
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14 (Accessed 2021-11-10). 
